---
title: "Assignment"
author: "Week 5"
date: "10/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

In this assignment we'll consider the scaling analysis from the 
slides: the New York Times German and US party scaling at the end.

For convenience, computational and otherwise, we'll use the `ca` 
package to do our scaling. For reference, this is the least squares
version of the model shown in the slides and consequently tends 
to give very similar results but rather quicker. So now is a good 
time to install the `ca` package (do it at the console, just once, not in this document).

Our data will be the results of a long running large scale 
human text analysis project once called the Comparative 
Manifestos Project (CMP) and subsequently various other things
that those of us who work in this area have forgotten. So we'll 
call it the CMP data. More about it can be found over here at the WZB: https://manifesto-project.wzb.eu

In short, lots of coders identify policy assertions in 
the platforms / manifestos / statements of policy preference 
of parties across Europe, North and South American, and a few other places, and code them into one of about 56 categories
(this changes over time but there are always a core 56). It's a 
histprical record and generates count data of the kind that a 
topic model or sentence classifier might automate, e.g. in 2002 the German Free Democrat Party had 14 sentences assigned to policy category `101` (Foreign Special Relationships: Positive) 
out of 1982 that were codable.

The gory details of the categories, parties, etc. can be found 
in the pdf codebook in the `data` folder which you can peruse 
at your leisure, but need not detain us now.

This kind of count data is also exactly the same sort of data as would come out of dictionary based content analysis, or if we had counted words rather than categories or topics. Consequently it can be scaled just the same 
way as we scaled the debate to infer speaker/document positions in class.

We'll start by loading the data set, correcting a small coding 
mistake, and turning the CMP's percentages back into the counts 
they were originally so we can treat it like the document 
feature matrix that it fundamentally is.

But before we begin, a quick note: you should feel free to get help if you 
are having trouble figuring out what the code is doing or how to adjust it.
Lightly adjusting the existing code is quite sufficient for this assignment, 
although impressive own coding is much appreciated.

## The data

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggrepel) 
library(ca)
library(dplyr)
library(quanteda)
library(ggplot2)

theme_set(theme_minimal())
```



```{r,message=FALSE, warning=FALSE}
# Read in, ignoring the parse failures because they're on a
# variable we won't use
allcmp20 <- read_csv("data/MPDataset_MPDS2020a.csv")
```

Now to give a nice abbreviation to the Pirates. It should 
probable be "ARGH", or "Avast!" or something, but we'll just make it "Pirates"
```{r}
allcmp20$partyabbrev[allcmp20$partyname == "Pirates"] <- "Pirates"
```
Now we'll re-inflate the counts and throw away all the electoral 
and other info in the data that we won't use
```{r}
cmp20 <- allcmp20 %>%
  select(edate, countryname, partyname, partyabbrev, total, peruncod,
         matches("per\\d\\d\\d$")) %>% # per%d%d%d% are subcategories, so we ignore them
  mutate(edate = as.Date(edate, format = "%d/%m/%Y"),
         eyear = lubridate::year(edate), # make a nice year just in case we want to filter with it
         label = paste(partyabbrev, eyear, sep = ":"), # for graphing
         across(starts_with("per"), function(x) round(total * (x/100)))) %>% # inflate the counts
  rename(uncoded = peruncod) # and rename the uncoded sentence counts
```
Phew. That was a fairly typical bit of data cleaning code, which 
takes up a lot of data science-ing time. Study it a little 
if you think this is in your future. But we'll just use it 
below.

The column `per101` is now the *count* of sentences coded into category 101. So, as noted above: 14 for the FDP in 2002. 
Feel free to confirm that.

## The metadata

Now to pull in the labels corresponding to the meaning of each category. This is metadata, since it is a mapping from the obscure
column header codes to two facts about each code.
```{r,message=FALSE}
itemcodes <- read_csv("data/itemcodes.csv")
head(itemcodes)

rile_right <- itemcodes$name[itemcodes$rile.valence == 1]
rile_left <- itemcodes$name[itemcodes$rile.valence == -1]
rile_none <- itemcodes$name[itemcodes$rile.valence == 0]

head(rile_right)
```
Here we've got a table of basic information about the CMP's coding 
scheme and also a "rile.valence". In this data there is a core set 
of categories that the authors believe is the basis of left-right
ideology measures. The categories in `rile_right` are the 
categories whose emphasis indicates a more right position
and `rile_left` a more left position. The `rile_none` categories
are just the remainder that are coded, but not supposed to 
be as important (or something. The codebook elaborates).
We'll use them later.

## Analysis: Germany

This is historical data, but we'll start our analysis somewhat 
arbitrarily at the beginning of the century.
```{r}
oldest <- as.Date("2000-01-01")
```
and focus on Germany.
```{r}
de <- filter(cmp20, edate > oldest, countryname == "Germany")

de[1:3, 1:7]
```
Because we're not in `quanteda` we'll have to manage the 
"docvars" and the document feature matrix info separately,
so let's first pull out the dfm-like part: the counts
```{r}
de_counts <- select(de, starts_with("per"))

de_counts[1:3, 1:7] # quick peek at the top left corner
```
and switch into a old-style `data.frame` representation that has
 rownames and colnames like you remember before you knew anything 
 about the tidyverse.
```{r}
de_mat <- data.matrix(de_counts)
rownames(de_mat) <- de$label
colnames(de_mat) <- itemcodes$name

de_mat[1:3,1:2] # top left corner of our dfm equivalent
```
Looks like dfm, right? If we've got this kind of object available then 
we're ready to scale.

## Model 

Now to scale these counts and see what they tell us about parties.
We'll use `ca` rather than `wordfish` because we'd like to 
be able to efficiently scale in multiple dimensions. So let's load 
the package and run the scaling function, conveniently called 
`ca` also.
```{r}
library(ca)

mod1 <- ca(de_mat)
```
`mod1` has a lot of stuff in it. Of primary interest to use
are the document positions "theta" and the word positions "beta".
CA calls these row coordinates and column coordinates respectively
(which makes sense) and these are tucked inside the model
in matrix form. Let's get them out and take a look.
```{r}
betas <- mod1$colcoord
thetas <- mod1$rowcoord

dim(thetas) # documents by (surprisingly large numbers of) dimensions
dim(betas) # features / policy categories by dimensions
```
Note: the positions of the i-th party in an election on the j-th dimension is the (i,j)-th element of `thetas`

Now it's your turn. Extract the positions for the parties on the first dimension and sort the results. Do these positions seem 
to make sense as ideological positions? If you don't have intuitions
about German parties, borrow a nearby German (there are plenty
locally) and ask them.
```{r}
#extracting a data frame for thetas/positions from a thetas matrix
party_positions <- data.frame(thetas[x=1:28,y=1]) %>% 
  #renaming to positions
  select(position=thetas.x...1.28..y...1.) %>% 
  #sorting results in a descending order using the arrange function
  arrange(desc(position)) %>% 
  #transforming the resultant matrix rows to dataframe columns
  rownames_to_column(var = "label") %>% 
  #stacking the theta names to the data frame
  data.frame(parameter = rep("theta", nrow(thetas))) %>% 
  #rearranging the columns
  select(label, parameter, position)
  
```
Interpretation
```
These positions seem to make sense since from the scores we are seeing right wing parties such as AFD, FDP and CDU falling with strong positive scores while others lean towards the left wing.
```

Now make a histogram of these positions, and provide a one sentence 
summary of what you see.
```{r}
#plotting with ggplot2
ggplot(party_positions, aes(position, color = parameter, label = label)) +
  #choosing histogram as my geom and specifying 30 bins
  geom_histogram(bins = 30)+
  #attaching text to the histogram and adjusting the size of the text to fit the plot
  geom_text_repel(data = party_positions, y=2, col="black", size=3)+
  #omotting grey color to make the plot brighter
  theme_light()
```

Summary:
```
Following a normal distribution, right wing parties are distributed to positive scores while left wing ones are distributed to negative scores.
```
Now do the same for the `betas` on dimension 1:
```{r}
pos_betas <- data.frame(betas[x=1:56,y=1]) %>% 
  #renaming to position column
  select(position=betas.x...1.56..y...1.) %>% 
  #sorting in a descending order using the arrange function
  arrange(desc(position)) %>% 
  #transforming the resultant matrix rows to dataframe columns
  rownames_to_column(var = "label") %>% 
  #replicating the beta names to the data frame
  data.frame(parameter = rep("beta", nrow(betas))) %>% 
  #rearranging the columns
  select(label, parameter, position)
  
```

Plotting a histogram for betas
```{r}
#plotting with ggplot2
ggplot(pos_betas, aes(position, color = parameter, label = label)) +
  #choosing histogram as my geom and specifying 30 bins
  geom_histogram(bins = 60)+
  #attaching text to the histogram and adjusting the size of the text to fit the plot
  geom_text_repel(data = pos_betas, y=3, col="black", size=3)+
  #omitting grey color to make the plot brighter
  theme_light()
```

Summary:
```
Right wing ideologies such as external security, free entreprise and incentives are skewed to the positive side while left wing ones such as anti-imperialism are in the negative ones.
```

Focusing now on the `betas` can you assign a rough interpretation 
to the second and third dimensions on the basis of how they 
order?
```{r}
#for the second dimension
beta_sec_dim <- data.frame(betas[x=1:56,y=2])%>% 
  #renaming to position column
  select(position=betas.x...1.56..y...2.) %>% 
  #sorting in a descending order using the arrange function
  arrange(desc(position)) %>% 
  #transforming the resultant matrix rows to dataframe columns
  rownames_to_column(var = "label") %>% 
  #replicating the beta names to the data frame
  data.frame(parameter = rep("beta", nrow(betas))) %>% 
  #rearranging the columns
  select(label, parameter, position)


#for third dimension
beta_third_dim <- data.frame(betas[x=1:56,y=3])%>% 
  #renaming to position column
  select(position=betas.x...1.56..y...3.) %>% 
  #sorting in a descending order using the arrange function
  arrange(desc(position)) %>% 
  #transforming the resultant matrix rows to dataframe columns
  rownames_to_column(var = "label") %>% 
  #replicating the beta names to the data frame
  data.frame(parameter = rep("beta", nrow(betas))) %>% 
  #rearranging the columns
  select(label, parameter, position)

```
Interpretation:
```
The second dimension suggest that ideologies contained in the spectrum have to do with populism. This is because ideologies such as Internationalism: Negative, Multiculturalism: Negative, National Way of Life: Positive and Protectionism: Positive score highly since they are being emphasized while dimensions such as Economic Goals score negative.

The third dimension spectrum has to do with foregn policy since in this spectrum, ideologies such as Internationalism: Negative, European Community/Union: Negative, Political Corruption and Protectionism: Positive score highly.
```

## Graphics

Time for a picture. Let's plot the parties and the policy categories
together in the first two dimensions. It will be helpful to 
make a data.frame that will make plotting easy for ggplot.
```{r}
# stack the betas and thetas with a label noting 
# which row is which, and a label for the graphic
plotdata <- data.frame(parameter = c(rep("beta", nrow(betas)),
                                     rep("theta", nrow(thetas))),
                       rbind(betas[,1:2], thetas[,1:2])) %>%
  rownames_to_column(var = "label")

head(plotdata)

```
Now we can plot them
```{r, fig.height = 10, fig.width = 10}
ggplot(plotdata, aes(Dim1, Dim2, color = parameter, label = label)) +
  geom_point() +
  geom_text_repel() +
  scale_colour_manual(values = list(beta="grey", theta = "black"),
                      guide = FALSE)

```

Oof. That's quite busy.

Let's try only plotting the `rile_left` and `rile_right` categories. 
We'll do this by filtering out all the `rile_none` rows of the data 
and using the same plot code.
```{r, fig.height = 10, fig.width = 10}
filter(plotdata, !(label %in% rile_none)) %>%
  ggplot(aes(Dim1, Dim2, color = parameter, label = label)) +
  geom_point() +
  geom_text_repel() +
  scale_colour_manual(values = list(beta="grey", theta = "black"),
                      guide = FALSE)
```

## Interpretation

Interpretation time. Consider the SPD manifesto in the 2017 election. According 
to this scaling model, does it emphasize each of the following themes *more* than other parties, *less* than other parties, or about the *same*?

- Political Authority
- Law and Order
- Constitutionalism
- Controlled Economy

Hint: revisit the final parts of the prerecorded video.

Interpretation:
```
- Political Authority: It emphasizes less of it/tends not to talk about it since its coordinates are in the opposite of where the SPD 2017 manifesto falls.

- Law and Order: It neither emphasizes it more or less than usual since its coordinates are located 90 degrees from the SPD 2017 manifesto.

- Constitutionalism: It is emphasized at the same rate since its coordinates are located in the middle from where SPD 2017 manifesto falls.

- Controlled Economy: It uses it strongly since it falls in the same quadrant with where coordinates for SPD 2017 manifesto falls.
```

## Time Series plots

We've been collapsing the temporal dimension here. Let's add those nicely formatted
dates we made earlier to the data and see how the parties look over time in terms of
the first scaled dimension
```{r}
ts_plotdata <- data.frame(date = de$edate, 
                          party = de$partyabbrev,
                          position = thetas[,1])
```
How about you make the positions over time plot and describe qualitatively what you see
```{r}
#using ggplot to plot from the give data
ggplot(ts_plotdata, aes(x= position, y= date, label = party))+
  #choosing points as my geoms
  geom_point()+
  #attaching text to the plot
  geom_text_repel(data=ts_plotdata, col="blue")+
  #specifying color fills
  scale_color_manual(values = list(position = "black"), guide = FALSE)+
  #omitting the grey color to make the plot brighter
  theme_light()
```

Interpretation:
```
I see how profiles of parties namely SPD, Greens, LINKE, FDP and CDU/CSU have been shifting from right to left, atleast in economic dimensions. Conversly, AFD has strikingly progressed to far right.
```

## Projection

Following the NYT piece, let's (perhaps unwisely) ask the question, where the two main 
US parties would be if they emphasized what they really do emphasize in their
platforms but were somehow translated into German and found themselves 
competing in a Federal Election. 

For this, we'll fit the model to the regular German parties and the project 
the US parties in based on the German `betas`. We can do this conveniently
by giving the `ca` function the row numbers of the US parties and asking it
to treat these as 'supplementary', that is, to define the scaled space without 
them and then to place them in it.

Most of this code follows the code above, but watch out for the differences
```{r}
usde <- filter(cmp20, edate > oldest,
               countryname %in% c("Germany", "United States"))
usde_counts <- select(usde, starts_with("per"))

usde_mat <- data.matrix(usde_counts)
rownames(usde_mat) <- usde$label
colnames(usde_mat) <- itemcodes$name

# which indices are to be projected?
extras <- which(usde$countryname == "United States")
extras

# now fit the model with supplementary rows 
mod2 <- ca(usde_mat, suprow = extras)

# and extract the positions
usde_betas <- mod2$colcoord
usde_thetas <- mod2$rowcoord
```
This time you plot it. Note: You *may* find it easier to read if you filter out *all* the betas, not just the `rile_none` ones.

```{r, fig.height = 10, fig.width = 10}
#stacking betas and thetas with label noting and specifying rows for the graphic
plotdata_usde <- data.frame(parameter = c(rep("beta", nrow(usde_betas)),
                                          rep("theta", nrow(usde_thetas))),
                            rbind(usde_betas[,1:2], usde_thetas[,1:2])) %>% 
  rownames_to_column(var = "label") %>% 
  #filtering out betas to remain with thetas
  filter(!(label %in% c(rile_none, rile_left, rile_right)))

#plotting with ggplot2
ggplot(plotdata_usde, aes(Dim1, Dim2, color = parameter, label = label)) +
  #choosing point as our geoms
  geom_point() +
  #specifying colors for text in the plot
  geom_text_repel(data = plotdata_usde, col="dark blue") +
  #specifying color for the dots
  scale_colour_manual(values = list(theta = "purple"),
                      guide = FALSE)+
  #omotting grey color to make the plot brighter
  theme_light()
```

Provide a brief qualitative description of what you see.

Interpretation:
```
The evolution of party platforms yearly convey the following
-Democrats and SPD emphasize on similar topics/ideologies since their coordinates are stuck around the same area. In the same light, Republicans and AFD emphasize on similar ideologies.

-Republicans appear to be far right while Democrats are center right.
```

## A more sensible projection?

Assuming the projection question makes sense (it probably does better for 
some pairs of countries than others, e.g. ones that share an electoral 
system or some relevant history), now project a *different*, perhaps more reasonable comparison country into the German space and tell us very briefly what you see and 
whether you think it makes sense.

```{r, fig.height = 10, fig.width = 10}
#projecting Austria that shares an electroal system with Germany
#filtering Germany and Austria
atde <- filter(cmp20, edate > oldest,
               countryname %in% c("Germany", "Austria"))
#filtering all the counts
atde_counts <- select(atde, starts_with("per"))
#generating a matrix
atde_mat <- data.matrix(atde_counts)
#giving names to rows
rownames(atde_mat) <- atde$label
#giving names to columns
colnames(atde_mat) <- itemcodes$name

#selecting indices to be projected
extras_at <- which(atde$countryname == "Austria")
extras_at

#fitting the model with supplementary rows
mod3 <- ca(atde_mat, suprow = extras_at)

#extracting betas (columns) from the fitted model (mod3)
atde_betas <- mod3$colcoord
#extracting thetas (rows) from the fitted model (mod3)
atde_thetas <- mod3$rowcoord


#preparing data for plotting
#stacking betas and thetas with label noting and specifying for the graphic
plotdata_atde <- data.frame(parameter = c(rep("beta", nrow(atde_betas)),
                                          rep("theta", nrow(atde_thetas))),
                            rbind(atde_betas[,1:2], atde_thetas[,1:2])) %>%
  #transforming matrix rows to a data frame rows
  rownames_to_column(var = "label") %>% 
  #filtering out betas to remain with thetas
  filter(!(label %in% c(rile_none, rile_left, rile_right)))


#plotting with ggplot2
ggplot(plotdata_atde, aes(Dim1, Dim2, color = parameter, label = label)) +
  #choosing dots as my geoms
  geom_point() +
  #specifying colors for text in the plot
  geom_text_repel(data = plotdata_atde, col="navy blue") +
  #specifying color for the dots
  scale_colour_manual(values = list(theta = "chocolate"),
                      guide = FALSE)+
  #omotting the grey background to make the plot brighter
  theme_light()
```
Interpretation:
```
FPO and AFD appears to be emphasizing similar ideologies of far right. SPO and SPD emphasizes similar ideologies of center right. Grune and Greens emphasize similar ideologies of far left.
```

## Finally

What do you think is the difference making a space using both countries 
and making a space with one and projecting another into it? Under what 
circumstances would these differ?

Interpretation:
```
Making a space using both countries returns projections that are more spread out between countries - there is bigger gaps from point to point. A good example is the US and Germany. But projecting one country into another one retuns projections that are not that spread out with smaller gaps. A good example is Germany and Austria. 

Also projecting using both countries leads to a distribution that is skewed to the left while projecting one country into another leads to a distribution that is more or less evenly distributed.

These differ under the following circumstances
-There is a comparison of a system with more than four parties against one with only two parties
-Variations in averages projecting to the dimensions from the center(0,0) which is the average of averages.
```



